---
title: "Tidy Rain cloud plot"
author: "Paula Andrea Martinez @orchid00"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I loved the blog post for [raincloud plots](https://micahallen.org/2018/03/15/introducing-raincloud-plots/). 
On a rainy day I've tried to tidy the code. 

### 1. Libraries
```{r}
library(tidyverse)
```

### 2. source geom_flat_violin
```{r}
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

### 3. Read in data
```{r}
my_data <- read_csv("https://data.bris.ac.uk/datasets/112g2vkxomjoo1l26vjmvnlexj/2016.08.14_AnxietyPaper_Data%20Sheet.csv")
#colnames(my_data)
#glimpse(my_data)
```

### 4. Tidy data to long format
```{r}
my_datal <- my_data %>% 
  select(Participant, AngerUH, DisgustUH, FearUH, HappyUH) %>% 
  gather(c("AngerUH", "DisgustUH", "FearUH", "HappyUH"), 
         key = "EmotionCondition", value = "Sensitivity")

head(my_datal)
```

### 5. Make a new theme
```{r}
raincloud_theme <- theme(
  text = element_text(size = 10),
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text = element_text(size = 14),
  axis.text.x = element_text(angle = 45, vjust = 0.5),
  legend.title = element_text(size = 16),
  legend.text = element_text(size = 16),
  legend.position = "right",
  plot.title = element_text(lineheight = .8, face = "bold", size = 16),
  panel.border = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_blank(),
  axis.line.x = element_line(colour = "black", size = 0.5, linetype = "solid"),
  axis.line.y = element_line(colour = "black", size = 0.5, linetype = "solid"))
```

### 6. Calculate summary stats 
```{r}
lb <- function(x) mean(x) - sd(x)
ub <- function(x) mean(x) + sd(x)

sumld <- my_datal %>% 
  select(-Participant) %>% 
  group_by(EmotionCondition) %>% 
  summarise_all(funs(mean, median, lower = lb, upper = ub))

sumld
groups(sumld) # just to show that after summarise there is no groups
```

### 7. Ready to plot, first raincloud plot! 
```{r}
g <- 
  ggplot(data = my_datal, 
         aes(x = EmotionCondition, y = Sensitivity, fill = EmotionCondition)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8) +
  geom_point(aes(y = Sensitivity, color = EmotionCondition), 
             position = position_jitter(width = .15), size = .5, alpha = 0.8) +
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) +
  scale_color_brewer(palette = "Spectral") +
  scale_fill_brewer(palette = "Spectral") +
  coord_flip() + # flip or not
  theme_bw() +
  raincloud_theme

g
```

### 8. Same plot - replacing the boxplot with a mean and confidence interval 
```{r}
g <- 
  ggplot(data = my_datal, 
         aes(x = EmotionCondition, y = Sensitivity, fill = EmotionCondition)) +
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8) +
  geom_point(aes(y = Sensitivity, color = EmotionCondition), 
             position = position_jitter(width = .15), size = .5, alpha = 0.8) +
  geom_point(data = sumld, aes(x = EmotionCondition, y = mean), 
             position = position_nudge(x = 0.3), size = 2.5) +
  geom_errorbar(data = sumld, aes(ymin = lower, ymax = upper, y = mean), 
                position = position_nudge(x = 0.3), width = 0) +
  expand_limits(x = 5.25) +
  guides(fill = FALSE) +
  guides(color = FALSE) +
  coord_flip() + # flip or not?
  scale_color_brewer(palette = "Spectral") +
  scale_fill_brewer(palette = "Spectral") +
  theme_bw() +
  raincloud_theme

g
# note: this gives a Warning: Ignoring unknown aesthetics: y
# It is from the geom_errorbar, but when removed gives and error
# Error in FUN(X[[i]], ...) : object 'Sensitivity' not found

```

You can find the [R markdown file here](https://github.com/orchid00/R4da/blob/master/ggplot2_tricks/tidy_raincloudplot.Rmd) 
and the [R script here](https://github.com/orchid00/R4da/blob/master/ggplot2_tricks/tidy_raincloudplot.R).
Thanks [micahgallen](https://twitter.com/micahgallen) for posting beautiful 
plots!

A side joke, I managed to write twice the wrong name instead of rain cloud plots, 
rainbow plots. Maybe because of the colours provided in the plots, maybe because
I just don't like rainy days like today.