---
title: "Tidy tuesday - week 1 US Tuition Costs"
author: "Paula Andrea Martinez @orchid00"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidy tuesday
[Tidy tuesday](https://github.com/rfordatascience/tidytuesday) seemed like a really nice game for me, hence I've tried to do something quick with the dataset of the week.

## 1. Libraries
```{r}
library(tidyverse)
library(readxl)
library(lubridate)
```

## 2. download the dataset
The dataset has been downloaded and stored in a `data` folder of the project.
```{r}
download.file(url = "https://github.com/rfordatascience/tidytuesday/blob/master/data/us_avg_tuition.xlsx?raw=true", destfile = "../data/us_avg_tuition.xlsx")
```

## 3. Read the data
```{r}
raw <- readxl::read_excel("../data/us_avg_tuition.xlsx")
```

## 4. Explore the data a bit
```{r}
class(raw)
```

```{r}
glimpse(raw)
```

There are twelve year observations per state. I decided to use all the years.

## 5. tidy the data
```{r}
# tidy dataset
mydata <- gather(raw, key = Year, value = Tuition, -State)
glimpse(mydata)
```
```{r}
# clean years to have for digit numbers
mydata <- 
    mydata %>% 
    mutate(Year = str_sub(Year, start = 1, end = 4))

glimpse(mydata)
```

### 5.1 How many states are in this dataset?

```{r}
unique(mydata$State)
```

## 6. Mean by state
```{r}
mydata1 <-
    mydata %>% 
    select(State, Tuition) %>% 
    group_by(State) %>% 
    summarise(Mean_State = mean(Tuition)) %>% 
    arrange(desc(Mean_State))
glimpse(mydata1)
```

```{r include=FALSE}
# Setting a custom ggplot2 function ---
# *** Functional Programming ***
# This function makes a pretty ggplot theme
# This function takes no arguments!
theme_LPD <- function(){
  theme_bw() +
    theme(axis.text.x = element_text(size = 12, vjust = 1, hjust = 1),
          axis.text.y = element_text(size = 12),
          axis.title.x = element_text(size = 14, face = "plain"),             
          axis.title.y = element_text(size = 14, face = "plain"),             
          panel.grid.major.x = element_blank(),                                          
          panel.grid.minor.x = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.y = element_blank(),  
          plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), units = , "cm"),
          plot.title = element_text(size = 20, vjust = 1, hjust = 0.5),
          legend.text = element_text(size = 12, face = "italic"),          
          legend.title = element_blank(),                              
          legend.position = c(0.8, 0.8))
}
```

## 7. Visualising the mean tuition by state
```{r fig.height=10}
ggplot(mydata1) +
    geom_bar(aes(x = reorder(State, Mean_State), y = Mean_State),
             stat = "identity", position = "dodge",
                 fill = "cadetblue3") +
    coord_flip() +
    theme_LPD()
    
```

## 8. Another way of visualizing in respect of the mean
```{r}
mydata2 <-
    mydata %>% 
    group_by(State) %>% 
    mutate(Mean_State = mean(Tuition)) %>% 
    ungroup() %>% 
    mutate(AwayFromMean = Mean_State - mean(Mean_State)) %>% 
    mutate(Category = ifelse(AwayFromMean < 0, "less than the mean", "higher than the mean")) %>% 
    arrange(desc(AwayFromMean)) %>% 
    select(State, AwayFromMean, Category) %>% 
    distinct()
glimpse(mydata2)
```


## 9. Which states have higher and lower tuition from the mean?
```{r fig.height=10}
ggplot(mydata2) +
    geom_bar(aes(x = reorder(State, -AwayFromMean), y = AwayFromMean, 
                 fill = Category),
             stat = "identity", position = "dodge") +
    geom_hline(yintercept = 0, linetype = 2) +
    coord_flip() +
    scale_fill_brewer() +
    labs(x = "State", y = "Away from the mean tuition") +
    theme_LPD()
    
```


This is an Rmarkdown document, you can find the code [here](https://github.com/orchid00/R4da/blob/master/ggplot2_tricks/us_avg_tuition.Rmd).