---
title: "Motivation R and graphics"
author: "Paula Andrea Martinez"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
library(ggplot2)
```

The **g**rammar of **g**raphics is a language for describing graphs.
Here are some common examples from the package ggplot2 to inspire you on how to use R to show your data.

```{r eval=FALSE}
install.packages("ggplot2")
library(ggplot2)
```


# Scatterplot

```{r fig.fullwidth=TRUE, cache=TRUE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```



# Colour full scatterplots

```{r fig.show='hold', fig.cap="Two plots in one figure environment"}
p <- ggplot(data = mpg, 
            mapping = aes(x = displ, y = hwy, 
                     colour = manufacturer)) + 
       geom_point()
p
p + geom_smooth(method = "lm")
```



# Stacked barplot

```{r fig.fullwidth=TRUE, cache=TRUE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, 
                         fill = clarity))
```



# Boxplots

```{r inlcude=FALSE}
library(datasets)

data(airquality)
airquality$Month <- factor(
             airquality$Month,
             labels = c("May", 
                        "Jun", 
                        "Jul", 
                        "Aug", 
                        "Sep"))
```

```{r fig.fullwidth=TRUE, warning=FALSE}
ggplot(data = airquality) + 
  geom_boxplot(mapping = 
                 aes(x = Month, y = Ozone))
```



# Polar plot
```{r fig.fullwidth=TRUE, cache=TRUE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut),
           width = 1) + 
  coord_polar()
```




# Open source

This handout was written in Rmarkdown, and uses the open-source [Tufte](https://rstudio.github.io/tufte/) style. It has been published on 
[Github pages](https://orchid00.github.io/ggplot2_examples.html) and also as a 
[PDF handout](https://github.com/orchid00/R4da/raw/master/scripts/ggplot2_examples.pdf). 

All of the information of my courses can be found on my Github repo R for Data Analysis https://github.com/orchid00/R4da. These resources are freely 
available under the Creative Commons - Attribution Licence. You may re-use and adapt the material in any way you wish, without asking permission, **provided you cite the original source**. That is a link back to the website [R for Data Analysis](https://github.com/orchid00/R4da) and my ORCID 0000-0002-8990-1985.

I acknowledge this publication is resulting from support of [Elixir-Belgium](https://www.elixir-belgium.org/) for my role as data science 
and bioinformatics trainer. 

Last update: `r Sys.Date()`
